---
title: "Practical Machine Learning Week 4 Project"
author: "ajibolafaki"
date: "June 19, 2018"
output: 
  md_document:
          variant: markdown_github
---

```{r setup, include=FALSE, warnings=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, the goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

## Load & Cleanse Data

```{r, results='hide', warnings=FALSE, message = FALSE}

library(caret)
library(tidyverse)
library(randomForest)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(RGtk2)
library(rattle)

trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training_df <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))
testing_df  <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))

# Exclude first seven variables because they are not predictors
#str(training_df)
training_df <- training_df[, -c(1:7)]

# Exclude columns with NAs from training dataset
training_df <- training_df[, colSums(is.na(training_df)) == 0]

# Check for near zero variance in training dataset
training_nzv <- nzv(training_df[,-ncol(training_df)],saveMetrics=TRUE)
#rownames(training_nzv)


# partition training dataset further into train and test data sets
inTrain  <- createDataPartition(training_df$classe, p=0.75, list=FALSE)
train_df <- training_df[inTrain, ]
test_df  <- training_df[-inTrain, ]

dim(train_df)
dim(test_df)

```

## Modeling : Decision Trees

```{r}
set.seed(21251)
fit_dt <- rpart(classe ~ ., data = train_df, method="class")
fancyRpartPlot(fit_dt)
predict_dt <- predict(fit_dt, test_df, type = "class")
confusionMatrix(predict_dt, test_df$classe)

```




## Modeling : Random Forest
In this section, we build a random forest model on the test set. We expect a small out of sample error estimate

```{r}
set.seed(21212)
fit_rf <- randomForest(classe ~ ., data = train_df, ntree = 1000)
predict_rf <- predict(fit_rf, test_df, type = "class")
confusionMatrix(predict_rf, test_df$classe)

```



## Modeling : GBM
In this section, we build a GBM model on the test set.

```{r results='hide', warnings=FALSE, message = FALSE}

set.seed(21237)
fit_gbm <- train(classe ~ ., data = train_df, 
                             trControl = trainControl(method = "repeatedcv", number = 5, repeats = 1),
                             method = "gbm")

```

```{r}

predict_gbm <- predict(fit_gbm, test_df)
confusionMatrix(predict_gbm, test_df$classe)

```

Since the random forest model performed better than GBM and Decision Trees, we run predicions on the 
test set using the random forest model.

```{r}

predict_rf_final <- predict(fit_rf, testing_df, type = "class")
predict_rf_final

```


